Practical seed-recovery for the PCG Pseudo-Random Number Generator

### 2020-08-19 article/main.tex
encoreremarques

\title{Predicting the PCG Pseudo-Random Number Generator In Practice}

@@

\begin{table}[htbp!]
  \centering
  \begin{tabular}{|c|Sc|Sc|c|}
	\hline
	n & $\vvvert H \vvvert \times \left\vvvert H^{-1} \right\vvvert$ & $ \lambda_1(\mathcal{L}) $ & minimum $t$ \\
	\hline
	3 & 2.87 & $4.09e^{12} \simeq 2^{41.9}$ & 26 \\
	4 & 2.06 & $2.44e^{14} \simeq 2^{47.8}$ & 20 \\
	5 & 3.77 & $1.72e^{15}\simeq 2^{50.6}$ & 18 \\
	6 & 2.69 & $1.03e^{16}\simeq 2^{53.2}$ & 15 \\
	\hline
  \end{tabular}
  \caption{minimal $t$ needed for a given $n$}
  \label{tab:babai}
\end{table}

@@

The heart of the algorithm is again the reconstruction of a truncated geometric
progression knowing the \(t=\ell+6\) upper bits. Looking at table~\ref{tab:babai}, we see that the best choice
consists in guessing 5 consecutive rotations and $\ell=14$ least-significant
bits. Therefore, \textsc{ReconstructPartialDifference} does $2^{57}$ iterations
of the inner loop, and succeeds deterministically.

@@

Knowing the values of $\Delta S_0[0:64+\ell]$ as well as the $\ell$
least-significant bits of $S_0$ and $c$ is sufficient to get rid the nastier
feature \textsf{PCG64}: armed with this knowledge, we can determine all the
subsequent rotations deterministically, at negligible cost, using
eq~\eqref{eq:find_rotation}. For each index $i$, it suffices to try the $64$
possible values of $r_i$; only one should satisfy
eq~\eqref{eq:find_rotation}. The complete pseudo-code is shown in
algorithm~\ref{algo:unknown_2}.

@@

Reconstructing $\Delta S_0$ from the $r_i$ is exactly the problem discussed in
section~\ref{sec:cvp_big}. This can be done by solving an instance of CVP in
dimension $63$. We use the off-the-shelf CVP solver embedded in \texttt{fplll}:
it runs in no measurable time.

>>>>

\title{Practical seed-recovery for the PCG Pseudo-Random Number Generator}

@@

\begin{table}[htbp!]
  \centering
  \begin{tabular}{|c|Sc|Sc|c|Sc|}
	\hline
	n & $\vvvert H \vvvert \times \left\vvvert H^{-1} \right\vvvert$ & $ \lambda_1(\mathcal{L}) $ & minimum $t$ & \(\left(1 + \vvvert H \vvvert \times \left\vvvert H^{-1} \right\vvvert \right) 2 \sqrt{n} 2^{64-t}\)\\
	\hline
	3 & 2.87 & $4.09e^{12} \simeq 2^{41.9}$ & 26 & \(3.69e^{12}\) \\
	4 & 2.06 & $2.44e^{14} \simeq 2^{47.8}$ & 20 &\(2.15e^{14}\)\\
	5 & 3.77 & $1.72e^{15}\simeq 2^{50.6}$ & 18 &\(1.5e^{15}\)\\
	6 & 2.69 & $1.03e^{16}\simeq 2^{53.2}$ & 15 &\(1.02e^{16}\)\\
	\hline
  \end{tabular}
  \caption{minimal $t$ needed for a given $n$}
  \label{tab:babai}
\end{table}

@@

The heart of the algorithm is again the reconstruction of a truncated geometric
progression knowing the \(t=\ell+6\) upper bits of four consecutive terms. Looking at table~\ref{tab:babai}, we see that the best choice
consists in guessing 5 consecutive rotations and $\ell=14$ least-significant
bits. Therefore, \textsc{ReconstructPartialDifference} does $2^{57}$ iterations
of the inner loop, and succeeds deterministically.

@@

Knowing the values of $\Delta S_0[0:64+\ell]$ as well as the $\ell$
least-significant bits of $S_0$ and $c$ is sufficient to get rid of the nastier
feature of \textsf{PCG64}: armed with this knowledge, we can determine all the
subsequent rotations deterministically, at negligible cost, using
eq~\eqref{eq:find_rotation}. For each index $i$, it suffices to try the $64$
possible values of $r_i$; only one should satisfy
eq~\eqref{eq:find_rotation}. The complete pseudo-code is shown in
algorithm~\ref{algo:unknown_2}.

@@

Reconstructing $\Delta S_0$ from the $r_i$ is exactly the problem discussed in
section~\ref{sec:cvp_big}. This can be done by solving an instance of CVP in
dimension $63$. We use the off-the-shelf CVP solver embedded in \texttt{fplll}:
it runs in negligible time.
